What do incoming university students believe about open science practices in

### 2022-03-29 analysis/Manuscript.Rmd
reduce words from Discussion, Abstract, and Introduction

---
title: What do incoming university students believe about open science practices in
  psychology?
author:
  - name: Jennifer L. Beaudry
    affiliation: '1,4'
    role:
      - Conceptualization
      - Data curation
      - Formal analysis
      - Investigation
      - Methodology
      - Project administration
      - Validation
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    corresponding: yes
    email: jen.beaudry@flinders.edu.au
    address: Research Development and Support, Flinders University, Bedford Park, 5042
  - name: Matt N. Williams
    affiliation: '2'
    role:
      - Investigation
      - Methodology
      - Validation
      - Writing - original draft
    email: m.n.williams@massey.ac.nz
  - name: Michael C. Philipp
    affiliation: '2'
    role:
      - Conceptualization
      - Investigation
      - Methodology
      - Project administration
      - Writing - original draft
    email: m.philipp@massey.ac.nz
  - name: Emily J. Kothe
    affiliation: '3'
    role:
      - Conceptualization
      - Data curation
      - Validation
      - Visualization
      - Writing - review & editing
    email: emily.kothe@deakin.edu.au
affiliation:
  - id: '1'
    institution: Flinders University, Australia
  - id: '2'
    institution: Massey University, New Zealand
  - id: '3'
    institution: Deakin University, Australia
  - id: '4'
    institution: Swinburne University of Technology, Australia
shorttitle: Psychology students' beliefs about open science
output:
  papaja::apa6_word: default
  papaja::apa6_pdf: default
acknowledgement: |
  The idea of this research was conceived at the 2018 UQ Open Science Conference and further developed at the 2019 Society for the Improvement of Psychological Science Unconference, What do first year psychology students think about (open) science?
  Our special thanks to those who contributed to the data collection effort including: Benjamin Le, Katie Gilligan-Lee, Annayah Prosser, Ian Fairholm, Jennie Ferrell, Kait Clark, Catherine Orr, Nick Burns, Rachel Searston, Rachel Stephens, Matthew Thompson, Aaron Drummond, Kathryn McGuigan, and Peter Cannon.
abstract: |
 Background: Understanding students’ naive conceptions about how science works and the norms that guide scientific best practice is important so that teachers can adapt to students’ existing understandings.
 Objective: We aimed to describe what incoming undergraduate students of psychology believe about reproducibility and open science practices.
 Method: We conducted an online survey with participants who were about to start their first course in psychology at a university (*N* = 239).
 Results: When asked to indicate how a researcher should conduct her study, most students endorsed several open science practices. When asked to estimate the proportion of published psychological studies that follow various open science practices, participants' estimates averaged near 50%. Only 18% of participants reported that they had heard the term "replication crisis".
 Conclusion: Despite media attention about the replication crisis, few incoming psychology students in our sample were familiar with the term. The students were nevertheless in favour of most open science practices, although they appeared to overestimate the prevalence of some of these practices in psychology.
 Teaching Implications: Teaching materials aimed at incoming psychology students should not assume pre-existing knowledge about open science or replicability.
keywords: open science, psychology, teaching, reproducibility, replication
wordcount: (need to update)
bibliography: ['r-references.bib','Undergrads & Open Science (public).bib']
floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
editor_options: 
  chunk_output_type: console
csl: "apa.csl" 
---

@@

These rapid changes present significant pedagogical challenges for the teacher of psychology. Keeping textbooks and other instructional materials up to date is difficult when supposedly well-established findings are contradicted by new replications emerging at a rapid pace. Furthermore, knowledge of emerging methodological practices is crucial for graduate student training. Even for students who do not go on to conduct research, an understanding of contemporary methodological practices---and problems with methodological practices in psychology---is essential for becoming informed and critical consumers of psychological knowledge. Studies have explored strategies for educating psychology students about replicability and open science practices [e.g., @chopikHowWhetherTeach2018; @graheHarnessingUndiscoveredResource2012; @jekelHowTeachOpen2020]. These initiatives may help ingrain open science norms and change attitudes about research practices, but we know little about what these students know or believe about open science research practices prior to entering the university classroom. This knowledge could be useful for two main reasons.

First, unlike some psychological phenomena in undergraduate courses (e.g., the internal workings of human senses), the replication crisis is frequently discussed in mainstream and social media [e.g., @ogradyPsychologyReplicationCrisis2020; @yongPsychologyReplicationCrisis2016; @yongPsychologyReplicationCrisis2018]. As such, students may have some knowledge about these issues prior to their formal studies. Teaching methods should be informed by some understanding of students' pre-existing levels of knowledge.

Second, anecdotal evidence suggests that students being taught about open science practices and reproducibility reforms (e.g., open sharing of data and analysis code) are sometimes surprised that these are not already standard practice. This could imply a need for educators to reinforce and build on students' “naive” impressions rather than radically altering their understanding of research practices. On the other hand, some evidence suggests that undergraduate students may quickly begin to engage in practices that hamper reproducibility. For example, @moranKnowItBad2021 found that 26.5% of the undergraduate students in their  sample admitted to “Conducting multiple statistical analyses on the same dataset in an attempt to find a statistically significant result” (p. 14; i.e., *p*-hacking), while 9.6% reported rounding down *p*-values to meet a significance threshold. Knowing the extent to which incoming psychology students may---or may not---be “naive open scientists” could help guide pedagogical approaches.

@@

# Methods
We preregistered the exploratory nature of this study at https://tinyurl.com/4p64mvwp [blinded]. Our materials, data, and analysis script are available at https://tinyurl.com/excm4heu [blinded]. Ethical approval for this study was granted by [blinded].

@@

Contacts from tertiary institutions in Australia, New Zealand, the UK, and the USA invited students enrolled in their first-year psychology course to participate in the online survey hosted on Qualtrics. These contacts were identified via several avenues, including a dedicated session at the Society for the Improvement of Psychological Science 2019 annual conference, a Twitter thread, and through contacts of the research team. Although we did not have inclusion criteria with respect to country, most contacts who advertised the study were from the United Kingdom, Australia, or New Zealand.

@@

We also measured participants' normative beliefs about how researchers actually conduct psychological science. We asked participants, "When you consider all of the psychology studies conducted globally each year, what percentage of them would have the following characteristics?" These practices corresponded approximately to our ten norm–counternorm pairs, with two exceptions (see Table \@ref(tab:normsinpracticeitems)). We excluded the critical thinking norm because it did not refer to externally observable behaviour, and we included another practice out of interest (data sharing on request). Participants indicated their estimate on a visual slider from 0% to 100%.

@@

Despite efforts to improve the replicability of research over the past decade, and corresponding media attention, most incoming psychology students in our sample were unaware of the replication crisis. Encouragingly, on the other hand, most tended to endorse open science norms more-so than counternorms. Notably this was not the case for all open science practices (namely sharing of pre-prints and open data). It may well be that scientific training has enshrined peer review as an essential practice that is robust to change.

Participants' estimates of the proportion of published psychological studies that apply each of the ten open science practices tended to hover around 50%. Although this could imply that participants literally believe these practices are roughly as commonly applied as not, it might also represent participants' uncertainty or indifference about the application of these practices in psychological studies. This said, 50% was the modal response for only a minority of practices (open access, data availability on request).

The fact that the mean estimated prevalence was no higher than 64% for any practice is inconsistent with the sometime-reported claim that incoming students tend to naively assume that science is usually conducted openly. Nevertheless, the students appear to have at least somewhat over-estimated the prevalence of some open science practices. For example, on average students estimated that nearly a third of published psychological studies share data openly and that nearly half of studies are preregistered. In contrast, an examination of a random sample of 250 psychology articles published between 2014 and 2017 [@hardwickeEstimatingPrevalenceTransparency2021] found that just 2% shared open data and 3% were preregistered.

## Strengths and Limitations
Despite a relatively large sample size, our results provide an imperfect picture of new psychology students’ methodological understandings. Our use of convenience samples coupled with the imprecision of polling students’ impressions at one point in time suggest that we can make only tentative generalisations to a wider population of incoming psychology students. In particular, the fact that our sample was relatively small and mostly comprised of students from Australia, NZ, or the United Kingdom limits the extent to which our findings can be generalised to other populations.

Our questions focused primarily on attitudes and beliefs so that we could better document the implicit understanding that guided students’ thoughts about the research process. However, we did not probe students' factual knowledge of most methodological findings or concepts, except the replication crisis. It is quite possible that reform in secondary school instruction has improved students’ understandings of more fundamental statistical (e.g., confidence interval estimation) and design (e.g., direct replications) considerations without necessarily influencing their open science attitudes and beliefs. Yet, research at the tertiary level casts doubt that the disciplinary reform from the past two decades has had much impact on the content of our teaching [@friedrichReplicatingNationalSurvey2018].

Although we have interpreted our findings with respect to the norm and counternorm items as reflecting  students' personal endorsement of practices, their perceptions of what we might perceive as the "right" answers could have also influenced their responses. Differences between endorsement of norms and counternorms may have also been affected by wording differences between the matched options; for example, the HARKing, information for replicability, and p-hacking counternorm items included the phrase "good scientific practice" whereas their matched norm items did not. More broadly, our choices of terminology in constructing items likely affected participants' responses to some degree. For example, the reference to keeping data "protected" in the open data counternorm item may have nudged participants towards endorsing this item, since failing to "protect" participants' data might sound irresponsible. Future research could examine the influence of revised language on participants' responses.

@@

## Conclusion and Teaching Implications
Overall these findings suggest that the incoming psychology students had a degree of sympathy for open science norms. Nevertheless, despite significant media attention on the replication crisis, few had heard of this crisis. Similarly, the students seemed to be relatively unfamiliar with responses to the replication crisis (e.g., open science practices). It is therefore important that teachers of psychology not assume pre-existing knowledge among incoming students.

For teachers interested in systematically surveying their students' existing understanding of open science practices in psychology, our survey questions could be a useful resource. These questions are available at https://tinyurl.com/excm4heu [blinded].

@@

Table A contains the correlations between norms and counternorms, with the 95% confidence intervals and the associated *p*-value.

@@

At a reviewer's request, we ran an additional exploratory analysis to determine whether
the proportion of students who are unaware of the replication crisis differs between psychology 
and non-psychology majors. We were able to code the majors for `r length(psyc_major$major_recoded)` students; of these `r round(length(psyc$crisis)/length(psyc_major$crisis)*100, 0)`% reported that psychology was their major.

Of those majoring in psychology, `r round(sum(psyc$crisis==1)/length(psyc$crisis)*100,0)`% indicated that they 
were aware of the replication crisis. Of those majoring in another program other than psychology, 
`r round(sum(not_psyc$crisis==1)/length(not_psyc$crisis)*100,0)`% indicated that they were aware of the replication crisis. A chi-square test of independence indicated that there was no significant difference between psychology
majors and non-majors in terms of their self-reported awareness of the replication crisis, X^2^ (`r chi$parameter`) = `r round(chi$statistic,2)`, *p* = `r round(chi$p.value,2)`. 
This lack of a signficant difference should be replicated in a larger population; however, it suggests that psychology instructors should not presume that students choosing to major in psychology are aware of the replication crisis.

>>>>

---
title: What do incoming university students believe about open science practices in
  psychology?
author:
  - name: Jennifer L. Beaudry
    affiliation: '1,4'
    role:
      - Conceptualization
      - Data curation
      - Formal analysis
      - Investigation
      - Methodology
      - Project administration
      - Validation
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    corresponding: yes
    email: jen.beaudry@flinders.edu.au
    address: Research Development and Support, Flinders University, Bedford Park, SA, 5042
  - name: Matt N. Williams
    affiliation: '2'
    role:
      - Investigation
      - Methodology
      - Validation
      - Writing - original draft
    email: m.n.williams@massey.ac.nz
  - name: Michael C. Philipp
    affiliation: '2'
    role:
      - Conceptualization
      - Investigation
      - Methodology
      - Project administration
      - Writing - original draft
    email: m.philipp@massey.ac.nz
  - name: Emily J. Kothe
    affiliation: '3'
    role:
      - Conceptualization
      - Data curation
      - Validation
      - Visualization
      - Writing - review & editing
    email: emily.kothe@deakin.edu.au
affiliation:
  - id: '1'
    institution: Flinders University, Australia
  - id: '2'
    institution: Massey University, New Zealand
  - id: '3'
    institution: Deakin University, Australia
  - id: '4'
    institution: Swinburne University of Technology, Australia
shorttitle: Psychology students' beliefs about open science
output:
  papaja::apa6_word: default
  papaja::apa6_pdf: default
acknowledgement: |
  The idea of this research was conceived at the 2018 UQ Open Science Conference and further developed at the 2019 Society for the Improvement of Psychological Science Unconference, What do first year psychology students think about (open) science?
  Our special thanks to those who contributed to the data collection effort including: Benjamin Le, Katie Gilligan-Lee, Annayah Prosser, Ian Fairholm, Jennie Ferrell, Kait Clark, Catherine Orr, Nick Burns, Rachel Searston, Rachel Stephens, Matthew Thompson, Aaron Drummond, Kathryn McGuigan, and Peter Cannon.
abstract: |
 Background: Understanding students’ naive conceptions about the norms that guide scientific best practice is important so that teachers can adapt to students’ existing understandings.
 Objective: We examined what incoming undergraduate students of psychology believe about reproducibility and open science practices.
 Method: We conducted an online survey with participants who were about to start their first course in psychology at a university (*N* = 239).
 Results: When asked to indicate how a researcher should conduct her study, most students endorsed several open science practices. When asked to estimate the proportion of published psychological studies that follow various open science practices, participants' estimates averaged near 50%. Only 18% of participants reported that they had heard the term "replication crisis".
 Conclusion: Despite media attention about the replication crisis, few incoming psychology students in our sample were familiar with the term. The students were nevertheless in favour of most open science practices, although they appeared to overestimate the prevalence of some of these practices in psychology.
 Teaching Implications: Teaching materials aimed at incoming psychology students should not assume pre-existing knowledge about open science or replicability.
keywords: open science, psychology, teaching, reproducibility, replication
wordcount: (need to update)
bibliography: ['r-references.bib','Undergrads & Open Science (public).bib']
floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
editor_options: 
  chunk_output_type: console
csl: "apa.csl" 
---

@@

These changes present significant pedagogical challenges for the teacher of psychology. Keeping textbooks and other instructional materials up to date is difficult when supposedly well-established findings are contradicted by new replications emerging at a rapid pace. Furthermore, knowledge of emerging methodological practices is crucial for graduate student training. Even for students who do not go on to conduct research, an understanding of contemporary methodological practices---and problematic methodological practices---is essential for becoming informed and critical consumers of psychological knowledge. Studies have explored strategies for educating psychology students about replicability and open science practices [e.g., @chopikHowWhetherTeach2018; @graheHarnessingUndiscoveredResource2012; @jekelHowTeachOpen2020]. These initiatives may help ingrain open science norms and change attitudes about research practices, but we know little about what students know or believe about open science research practices prior to entering the university classroom. This knowledge could be useful for two main reasons.

First, unlike some psychological phenomena in undergraduate courses (e.g., the internal workings of human senses), the replication crisis is frequently discussed in mainstream and social media [e.g., @ogradyPsychologyReplicationCrisis2020; @yongPsychologyReplicationCrisis2016; @yongPsychologyReplicationCrisis2018]. As such, incoming students might have some knowledge about these issues prior to their formal studies. Teaching methods should be informed by understanding students' pre-existing levels of knowledge.

Second, anecdotal evidence suggests that students being taught about open science practices and reproducibility reforms (e.g., open sharing of data and code) are sometimes surprised that these are not already standard practice. This could imply a need for educators to reinforce and build on students' “naive” impressions rather than radically altering their understanding of research practices. On the other hand, some evidence suggests that undergraduate students may quickly begin to engage in practices that hamper reproducibility. For example, @moranKnowItBad2021 found that 26.5% of the undergraduate students in their  sample admitted to “Conducting multiple statistical analyses on the same dataset in an attempt to find a statistically significant result” (p. 14; i.e., *p*-hacking), while 9.6% reported rounding down *p*-values to meet a significance threshold. Knowing the extent to which incoming psychology students may---or may not---be “naive open scientists” could help guide pedagogical approaches.

@@

# Methods
We preregistered the exploratory nature of this study at https://tinyurl.com/4p64mvwp [blinded]. Our materials, data, and analysis script are available at https://tinyurl.com/excm4heu [blinded]. Ethical approval was granted by [blinded].

@@

Contacts from tertiary institutions in Australia, New Zealand, the UK, and the USA invited students enrolled in their first-year psychology course to participate in the online survey hosted on Qualtrics. These contacts were identified via several avenues, including a dedicated session at the Society for the Improvement of Psychological Science 2019 annual conference, a Twitter thread, and through contacts of the research team. We did not have inclusion criteria with respect to country, but most contacts who advertised the study were from the United Kingdom, Australia, or New Zealand.

@@

We also measured participants' normative beliefs about how researchers actually conduct psychological science. We asked participants, "When you consider all of the psychology studies conducted globally each year, what percentage of them would have the following characteristics?" These practices corresponded approximately to our ten norm–counternorm pairs, with two exceptions (see Table \@ref(tab:normsinpracticeitems)). We excluded the critical thinking norm because it did not refer to externally observable behaviour, and we included another practice out of interest (data availability on request). Participants indicated their estimate on a visual slider from 0% to 100%.

@@

Despite efforts to improve the replicability of research and the corresponding media attention, most incoming psychology students in our sample were unaware of the replication crisis. Encouragingly, on the other hand, most tended to endorse open science norms more-so than counternorms. Notably this was not the case for all open science practices (namely sharing of pre-prints and open data). It may well be that scientific training has enshrined peer review as an essential practice that is robust to change.

Participants' estimates of the proportion of published psychological studies that apply open science practices tended to hover around 50%. Although this could imply that participants literally believe these practices are roughly as commonly applied as not, it might also represent participants' uncertainty or indifference about the application of these practices. This said, 50% was the modal response for only a minority of practices (open access, data availability on request).

That the mean estimated prevalence was no higher than 64% for any practice is inconsistent with the sometime-reported claim that incoming students tend to naively assume that science is conducted openly. Nevertheless, the students appear to have at least somewhat over-estimated the prevalence of some open science practices. For example, on average students estimated that nearly a third of published psychological studies share data openly and that nearly half of studies are preregistered. In contrast, an examination of a random sample of 250 psychology articles published between 2014 and 2017 [@hardwickeEstimatingPrevalenceTransparency2021] found that just 2% shared open data and 3% were preregistered.

## Strengths and Limitations
Despite a moderate sample size, our results provide an imperfect picture of new psychology students’ methodological understandings. Our use of convenience samples coupled with the imprecision of polling students’ impressions at one point in time suggest that we can make only tentative generalisations to a wider population of incoming psychology students. In particular, the fact that our sample mostly comprised students from Australia, NZ, or the United Kingdom limits the extent to which our findings can be generalised to other populations.

Our questions focused primarily on attitudes and beliefs so that we could better document the implicit understanding that guided students’ thoughts about the research process. However, we did not probe students' factual knowledge of most methodological findings or concepts, except the replication crisis. It is quite possible that reform in secondary school instruction has improved students’ understandings of more fundamental statistical (e.g., confidence interval estimation) and design (e.g., direct replications) considerations without necessarily influencing their open science attitudes and beliefs. Yet, research at the tertiary level casts doubt that the methodological reform in psychology has had much impact on the content of our teaching [@friedrichReplicatingNationalSurvey2018].

Although we have interpreted participants' responses to the norm and counternorm items as reflecting  students' personal endorsement of practices, their perceptions of what we might perceive as the "right" answers could have also influenced their responses. Differences between endorsement of norms and counternorms might have also been affected by wording differences between the matched options; for example, the HARKing, information for replicability, and *p*-hacking counternorm items included the phrase "good scientific practice" whereas their matched norm items did not. More broadly, our choices of terminology in constructing items likely affected participants' responses to some degree. For example, the reference to keeping data "protected" in the open data counternorm item may have nudged participants towards endorsing this item because failing to "protect" participants' data might sound irresponsible. Future research could examine the influence of revised language on participants' responses.

@@

## Conclusion and Teaching Implications
Our findings suggest that these incoming psychology students had a degree of sympathy for open science norms. Nevertheless, despite significant media attention, few had heard of the replication crisis. Similarly, the students seemed to be relatively unfamiliar with responses to the replication crisis (e.g., open science practices). It is therefore important that teachers of psychology not assume pre-existing knowledge among incoming students.

For teachers interested in systematically surveying their students' existing understanding of open science practices in psychology, our survey questions could be a useful resource (available at https://tinyurl.com/excm4heu [blinded]).

@@

Table A contains the correlations between norms and counternorms, with the 95% confidence intervals and associated *p*-values.

@@

At a reviewer's request, we ran an additional exploratory analysis to determine whether
the proportion of students who are unaware of the replication crisis differed between psychology 
and non-psychology majors. We were able to code the majors for `r length(psyc_major$major_recoded)` students; of these, `r round(length(psyc$crisis)/length(psyc_major$crisis)*100, 0)`% reported that psychology was their major.

Of those majoring in psychology, `r round(sum(psyc$crisis==1)/length(psyc$crisis)*100,0)`% indicated that they 
were aware of the replication crisis. Of those majoring in another program other than psychology, 
`r round(sum(not_psyc$crisis==1)/length(not_psyc$crisis)*100,0)`% indicated that they were aware of the replication crisis. A chi-square test of independence indicated that there was no significant difference between psychology
majors and non-majors in terms of their self-reported awareness of the replication crisis, X^2^ (`r chi$parameter`) = `r round(chi$statistic,2)`, *p* = `r round(chi$p.value,2)`. 
This lack of a significant difference should be replicated in a larger population; however, it suggests that psychology instructors should not presume that students choosing to major in psychology are aware of the replication crisis.
